Improving Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD Transient

### 2020-02-25 paper/cvpr2020/rebuttal/rebuttal.tex

\documentclass[10pt,twocolumn,letterpaper]{article}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{float}
% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{2224} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
\raggedbottom
\begin{document}
\title{Optimizing MDE with Global Depth Histogram
  Matching using a Single SPAD Transient}

\maketitle
\thispagestyle{empty}
%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
We thank our reviewers for their constructive feedback and will make every
effort to address their concerns.

We propose a technique for augmenting the capabilities of monocular depth
estimation using transient data from a single-pixel SPAD and diffuse laser
illumination using a straightforward histogram matching algorithm. We agree with
all of our reviewers that the idea of using such a depth transient is novel and
worth investigating.

\vspace{1mm}\noindent\textbf{Equivalence of scanning and diffuse SPAD (R1, R2)}
First, for
both scanned and diffuse modes, pileup can be corrected \cite{heide2018subpico} or avoided by operating in the
low-flux regime. 
Second, we empirically validate that the digitally summed scanned SPAD
measurements and the diffused measurements are almost identical. We will add
discussions of both of these points to the paper.
\vspace{-2mm}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/figure.pdf}
  \caption{Setup and captured transients for scanned and diffuse mode.}
\end{figure}

\vspace{-3mm}\noindent\textbf{Energy of diffuse SPAD (R1, R2, R3)}
Assuming an ambient spectral irradiance of $I_A = 0.1$ $mW/m^2$, we can calculate the
laser power required to achieve a minimum SBR of 5 for a diffuse scene at $r = 2$ m
and a field of view of $\theta = 40^\circ$ as
\begin{equation}
  P_{\text{min}} = I_A \cdot 4 r^2 \tan^2(\theta/2) \cdot SBR_{\text{min}}
\end{equation}
Which gives $P_{\text{min}} = 1.1$ W. laserpointersafety.com gives the Nominal
Ocular Hazard Distance for this laser as 0.34m, i.e. the laser light level is
safe beyond this distance. Given that there are VCELs with well above
this capability, we feel that our method is a very competitive, eye safe, option for short
range indoor scenes.

% McManamon 2012 provide the following equation for any LiDAR setup:
% \begin{equation}
%   P_R = P_T \cdot \frac{\rho_tA_t}{A_{\text{illum}}} \cdot \frac{A_{\text{rec}}}{\pi R^2} \cdot \eta_{\text{atm}}^2 \cdot \eta_{\text{sys}}
% \end{equation}
%   In the diffused setting, With a laser at 450 nm, and under the assumptions of $P_T=0.0003, \rho_t = 0.5,
%   A_t = A_{\text{illum}}, A_{\text{rec}} = 7\times 10^{-8}$ (corresponding to a
%   SPAD with aperture diameter $300$ $\mu$m), $R = 5, \eta_{\text{atm}} = 1$, and
%   $\eta_{\text{sys}} = 0.5$ we
%   accumulate $10^6$ signal photons in 0.04 sec, 0.8 sec in the low-flux regime.
%   At 10MHz, this corresponds to 8 million pulses, or the equivalent of scanning
%   each point in a $512 \times 512$ grid 30 times. (Include ambient flux levels?)

\vspace{1mm}\noindent\textbf{Justification for two-step method (R1)} We believe
our final architecture's strength is the fact that it is agnostic to the
MDE strategy used and will note this in the paper.

\vspace{1mm}\noindent\textbf{Kinect/Dual Pixels (R2)} While each of these methods has its own
benefits and disadvantages, our approach adds minimal additional hardware to a
single RGB camera. We will rephrase our comparisons to these methods accordingly.

\vspace{1mm}\noindent\textbf{Failure/limited improvement cases (R2)} We show at least one example in
the supplement where our method improves in RMSE by only 0.062. In general, our
method struggles with scenes where the inital depth map contains ordinal depth
errors.

\vspace{1mm}\noindent\textbf{Maximum SBR used for simulated results (R2)} The
experimental results are intended to show the performance of the method under
more realistic conditions. For example, the SBR of the outdoor scene was
apprximately 0.15. Detailed statistics for the experimental scenes will be
included to help strengthen our argument for practicality.

\vspace{1mm}\noindent\textbf{Interreflections (R2)} We neglect interreflections
as a second-order effect following existing work on flash LiDAR systems. \cite{iddan20013dstudio}

\vspace{1mm}\noindent\textbf{Uniform background assumption and negative
  histogram values (R2)} When operating in the low-flux condition, we follow the
many papers \cite{lindell2018} which model background photons as a constant term that adds equally to all bins.
Also, our algorithm clips all negative histogram values to zero after the background
subtraction step and we will update the section to reflect this.

\vspace{1mm}\noindent\textbf{Number of acquisition cycles (R1)} In simulation, the histogram
is directly computed via equation (2).
Scan times varied from 5 to 25 mins indoors and 3 mins for the outdoor
scene (the number of $512 \times 512$
frames can be computed from the scan time and the laser pulse rate of 10MHz).
While this seems excessively long, this was a deliberate decision to maximize the fidelity of our
ground truth reconstructions, not to simulate the real capture setting. We
will report the scan times and number of acquisitions in the paper.

\vspace{1mm}\noindent\textbf{Selection of $K$, $\ell$, and $u$ (R1)} For both simulation and
experiment, $K$ is chosen to tradeoff between speed and accuracy. For
$h_{\text{source}}$, $(\ell, u)$ are set to encompass all possible scene depths. For 
    $h_{\text{target}}$, $(\ell, u)$ are chosen based on the depth values for
    the bins $\hat n_{\text{first}}$ and $\hat n_{\text{last}}$, respectively.
    We will update the paper to clarify this point.

\vspace{1mm}\noindent\textbf{Use of overdramatic or imprecise language (R1, R2)} We will make various
wording adjustments, such as using ``moderate'' instead of
``slight'' to describe the improvement from median rescaling, using
``augmenting'' instead of ``optimizing'', and changing improper uses of
``histogram'' to ``transient''.

\vspace{1mm}\noindent\textbf{Teaser colormaps misleadingly scaled (R2)} We will adjust both
    figures to use the same scale.

\vspace{1mm}\noindent\textbf{Approximating Poisson dist. with Gaussian (R3)} We will add a
    footnote explaining our usage of this approximation where it occurs.

\vspace{1mm}\noindent\textbf{Typo in L388 (R3)} We will correct $n < n_{\text{last}}$ to $n <
    n_{\text{first}}$.

{\small
\bibliographystyle{ieeetr}
\bibliography{references}
}

\end{document}

### 2020-02-25 paper/eccv2020/main.tex
Adding eccv submission

% TODO:
%@article{caramazza2018neural,
%  title={Neural network identification of people hidden from view with a single-pixel, single-photon detector},
%  author={Caramazza, Piergiorgio and Boccolini, Alessandro and Buschek, Daniel and Hullin, Matthias and Higham, Catherine F and Henderson, Robert and Murray-Smith, Roderick and Faccio, Daniele},
%  journal={Scientific reports},
%  volume={8},
%  number={1},
%  pages={11945},
%  year={2018},
%  publisher={Nature Publishing Group}
%}
%

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\usepackage{float}
\usepackage[utf8x]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{dblfloatfix}

\input{mark_defs.tex}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{2224} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Optimizing Monocular Depth Estimation with Global Depth Histogram
  Matching using a Single SPAD Transient}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
Existing monocular depth estimation algorithms successfully predict the relative depth {order} of objects in a scene.
However, because of the fundamental scale ambiguity associated with monocular images, these algorithms fail at correctly predicting an object's true metric depth. 
In this work, we demonstrate how a depth histogram of the scene, which can be readily captured using a single-pixel diffused single-photon avalanche diode (SPAD), can be fused with the output of existing monocular depth estimation algorithms to resolve the depth ambiguity problem. 
We validate this novel sensor fusion technique experimentally and in extensive simulation. We show that it dramatically improves the performance of several state-of-the-art monocular depth estimation algorithms.
\end{abstract}

%%%%%%%%% BODY TEXT

@@

{\small
\bibliographystyle{ieeetr}
\bibliography{references}
}

\end{document}

>>>>

% updated April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016; AAS, 2020

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
% \usepackage{ruler}
% \usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{array}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{multirow}
\usepackage{float}
\usepackage[utf8x]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{dblfloatfix}

\input{mark_defs.tex}
% \DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
% \def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}

\def\eg{\emph{e.g.}} \def\Eg{\emph{E.g.}}
\def\ie{\emph{i.e.}} \def\Ie{\emph{I.e.}}
\def\cf{\emph{c.f.}} \def\Cf{\emph{C.f.}}
\def\etc{\emph{etc}} \def\vs{\emph{vs.}}
\def\wrt{w.r.t.} \def\dof{d.o.f.}
\def\etal{\emph{et al.}}
\graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
% Pages are numbered in submission mode, and unnumbered in camera-ready

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCVSubNumber{3668}  % Insert your submission number here

\title{Improving Monocular Depth Estimation with Global Depth Histogram
  Matching using a Single SPAD Transient} % Replace with your title

% INITIAL SUBMISSION 
%\begin{comment}
\titlerunning{ECCV-20 submission ID \ECCVSubNumber} 
\authorrunning{ECCV-20 submission ID \ECCVSubNumber} 
\author{Anonymous ECCV submission}
\institute{Paper ID \ECCVSubNumber}
%\end{comment}
%******************

% CAMERA READY SUBMISSION
\begin{comment}
\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
\end{comment}
%******************
\maketitle

\begin{abstract}
  Existing monocular depth estimation algorithms successfully predict the
  relative depth {order} of objects in a scene.
  However, because of the fundamental scale ambiguity associated with monocular
  images, these algorithms fail at correctly predicting an object's true metric
  depth.
  In this work, we demonstrate how a depth histogram of the scene, which can be
  readily captured using a single-pixel diffused single-photon avalanche diode
  (SPAD), can be fused with the output of existing monocular depth estimation
  algorithms to resolve the depth ambiguity problem.
  We validate this novel sensor fusion technique experimentally and in extensive
  simulation. We show that it dramatically improves the performance of several
  state-of-the-art monocular depth estimation algorithms.
  \dots
  \keywords{monocular depth estimation, single-photon avalanche diode, histogram
  matching, LiDAR}
\end{abstract}

@@

\clearpage
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references}
\end{document}

### 2020-02-25 paper/eccv2020/sections/simulation.tex
Adding eccv submission

To simulate a SPAD histogram, we take the provided depth map and 
calculate a weighted depth histogram by weighting the pixel contributions to
each depth bin by the luminance of each pixel. To model radiometric falloff, we
multiply each bin by $1/z^2$, and convolve with a modeled system temporal
response, which we approximate as a Gaussian with a full-width at half-maximum of 70~ps. We scale
the histogram by the total number of observed signal photon counts (set to 
$10^6$) and  add a fixed number of background photons $b \in \set{10^5, 2\times
10^4, 10^4}$. The background counts are evenly distributed across all bins to simulate the ambient and dark
count detections, and the different background levels correspond to
signal-to-background ratios (SBR) of $10, 50$ and $100$ respectively. Finally,
each bin is Poisson sampled to produce the final simulated SPAD transient.

% \textcolor{red}{Need more details: how did you simulate the SPAD measurements from this: discuss albedo, square distance falloff, Poisson noise, background, number of SPAD bins, define what SBR is etc.}

@@

We show an extensive quantitative evaluation in Table~\ref{tab:comparison}.
Here, we evaluate three recent monocular depth estimation techniques:
DORN~\cite{Fu2018}, DenseDepth~\cite{Alhashim2018}, and
MiDaS~\cite{Lasinger:2019}. To evaluate the quality of DORN and DenseDepth, we
report various standard error metrics. Moreover, we show a simple
post-processing step that rescales their outputs by the median ground truth
depth~\cite{Alhashim2018}. We also show the results of histogram matching the
output of the CNNs with the ground truth depth map histogram. Note that we do
not report the quality of the direct output of MiDaS as this algorithm does not
output metric depth. However, we do show its output histogram matched with the
ground truth depth map histogram. In all cases, post-processing the estimated
depth maps either with the median depth or depth histogram significantly
improves the absolute depth estimation, often by a large margin compared to the
raw output of the CNNs. Unfortunately, ground truth depth is typically not
accessible so neither of these two post-processing methods are viable in
practical application scenarios.

@@

In Figure~\ref{fig:results_simulated}, we also show qualitative results of our
simulations. For each of these scenes, we show the RGB reference image, the
ground truth depth map, the raw output of the DenseDepth CNN, the result of
rescaling the CNN output by the median ground truth depth, the result of
histogram-matching the CNN output by the ground truth depth map histogram, and
the result achieved by the proposed method for an SBR of 100. Error maps for all
the depth estimation methods are shown. As expected, the CNN outputs depth maps
that look reasonable but that have an average root mean squared error (RMSE) of
about 50--60~cm. Re-scaling this depth map by the median ground truth depth
value slightly improves the quality and histogram-matching with the ground
truth depth histogram shows a large amount of improvement. The quality of the
proposed method is close to using the oracle histogram, despite relying on
noisy SPAD measurements. Additional simulations using DenseDepth and other depth
estimation CNNs for a variety of scenes are shown in the supplement.

>>>>

To simulate a SPAD histogram, we take the provided depth map and 
calculate a weighted depth histogram by weighting the pixel contributions to
each depth bin by the luminance of each pixel. To model radiometric falloff, we
multiply each bin by $1/z^2$, and convolve with a modeled system temporal
response, which we approximate as a Gaussian with a full-width at half-maximum of 70~ps. We scale
the histogram by the total number of observed signal photon counts (set to 
$10^6$) and  add a fixed number of background photons $b \in \set{2\times 10^5, 10^5, 2\times
10^4, 10^4}$. The background counts are evenly distributed across all bins to simulate the ambient and dark
count detections, and the different background levels correspond to
signal-to-background ratios (SBR) of $5, 10, 50$ and $100$ respectively. Finally,
each bin is Poisson sampled to produce the final simulated SPAD transient.

@@

We show an extensive quantitative evaluation in Table~\ref{tab:comparison}.
Here, we evaluate three recent monocular depth estimation CNNs:
DORN~\cite{Fu2018}, DenseDepth~\cite{Alhashim2018}, and
MiDaS~\cite{Lasinger:2019}. To evaluate the quality of DORN and DenseDepth, we
report various standard error metrics~\cite{Eigen2014}. Moreover, we show a simple
post-processing step that rescales their outputs to match the median ground truth
depth~\cite{Alhashim2018}. We also show the results of histogram matching the
output of the CNNs with the ground truth depth map histogram. Note that we do
not report the quality of the direct output of MiDaS as this algorithm does not
output metric depth. However, we do show its output histogram matched with the
ground truth depth map histogram. In all cases, post-processing the estimated
depth maps either with the median depth or depth histogram significantly
improves the absolute depth estimation, often by a large margin compared to the
raw output of the CNNs. Unfortunately, ground truth depth is typically not
accessible so neither of these two post-processing methods are viable in
practical application scenarios.

@@

In Figure~\ref{fig:results_simulated}, we also show qualitative results of our
simulations. For each of these scenes, we show the RGB reference image, the
ground truth depth map, the raw output of the DenseDepth CNN, the result of
rescaling the CNN output with the median ground truth depth, the result of
histogram-matching the CNN output by the ground truth depth map histogram, and
the result achieved by the proposed method for an SBR of 100. Error maps for all
the depth estimation methods are shown. As expected, the CNN outputs depth maps
that look reasonable but that have an average root mean squared error (RMSE) of
about 50--60~cm. Rescaling this depth map to match the median ground truth depth
value slightly improves the quality and histogram-matching with the ground
truth depth histogram shows a large amount of improvement. The quality of the
proposed method is close to using the oracle histogram, despite relying on
noisy SPAD measurements. Additional simulations using DenseDepth and other depth
estimation CNNs for a variety of scenes are shown in the supplement.
